---
title: "damm-mcnip"
author: "Rose Z. Abramoff"
date: "July 14, 2015"
output: html_document
---
##DAMM-MCNiP is a C-N coupled microbial physiology model, created from the merger of DAMM (Davidson et al. 2012) and MCNiP (Finzi et al. 2015). 

##Model options
```{r options}
#runtime (hours)
times.def <<- seq(1,4560)
```

##Parameter options
```{r params}
#define parameters
#litter C:N should prob be higher than 27.6

pars <- list(e = 0.00026, cn_ex = 27.6, cn_s = 27.6, cn_m = 10, cn_enz = 3, p = 0.5, q = 0.5, a = 0.5, r = 0.008314, r_death = 0.00015, r_ecloss = 0.001, mic_to_soc = 0.5, mic_to_son = 0.5, doc_input = 0.0005, litter_c = 0.0005, don_input = 0.0005/27.6, litter_n = 0.0005/27.6, a_upt_c = 1.0815e11, vmax_0 = 1.0815E11, ea_upt_c = 61.77, ea_up = 61.77, km_upt_c = 0.3, km_upt_n = 0.3, km_c = 0.0025, km_n = 0.0025, cue = 0.31, km_o2 = 0.121, dgas = 1.67, o2airfrac = 0.209, bd = 0.8, pd = 2.52, frac = 0.000414, dliq = 3.17)

#specify min and max values of each parameter of interest
#used in: global sensitivity analysis, stability analysis
parRanges <- data.frame(min = c(30), max = c(100))
rownames(parRanges) <- c("ea_up")
pars2fit <- c(ea_up = 61.77)
```

##Pool options
```{r pools}
#pre-allocate pools
mic_c <- mic_n <- ec <- soc <- son <- doc <- don <- o2 <- vector(mode = "double", length = length(times.def)) 

#initialize pools
state <<- c(mic_c = 1.1957, mic_n = 0.1196, soc = 144.5986, son = 5.4413, doc = 0.00091631, don = 0.00049421, ec = 0.0325)

#specify pool(s) of interest in `sensvar`
#used in: global sensitivity analysis, local sensitivity analysis, parameter identifiability analysis
#intended use in parameter identifiability analysis: set `sensvar` = pools for which there are data, then assess collinearity to determine which parameter combinations require data
sensvar = c("soc")
```

##Fluxes
```{r fluxes}
#pre-allocate fluxes
decom_c <- decom_n <- upt_c <- upt_n <- enz_c <- enz_n <- eprod <- eloss <- growth_c <- growth_n <- growth <- death_c <- death_n <- overflow_c <- cmin <- nmin <- vector(mode = "double", length = length(times.def))
```

##Environmental Drivers
```{r envdr}
#Environmental inputs (190 days worth of data)
f <- read.csv("/Users/rzabramoff/Dropbox/bu/dissertation_research/damm-mcnip/hformat.csv", sep=",",header=T, na.strings="NA")
soilt <- f$SoilT #rep(13.73, length(times.def)) 
soilm <-  ifelse(f$SoilM > 0.5, 0.5, f$SoilM) #rep(0.3369, length(times.def))
porosity <- 1 - pars$bd/pars$pd;
```

##Load libraries
```{r loadlib}
library(deSolve)
library(FME)
```


##Model
```{r model}
#define model as a function
snap <- function(pars, times = times.def) {
derivs <- function(t, state, pars) {
   with(as.list(c(state, pars)), {
     #relevant model equations
     porosity = 1 - bd/pd; 
     o2 <- dgas * o2airfrac * (porosity - mois(t))^(4/3) #calculate [o2] at time t
     sol_soc <- dliq * mois(t)^3 * frac * soc #calculate available soc at reaction site
     sol_son <- dliq * mois(t)^3 * frac * son #calculate available son at reaction site
     vmax_upt_c <- vmax_upt_n <- a_upt_c * exp(-ea_upt_c / (r * (temp(t) + 273))) #calculate uptake kinetics
     vmax_c <- vmax_n <- vmax_0 * exp(-ea_up / (r * (temp(t) + 273))) #calculate depolymerization kinetics
     upt_c <- mic_c * vmax_upt_c * (doc / (km_upt_c + doc)) * (o2 / (km_o2 + o2)) #microbial C uptake
     cmin <- upt_c * (1 - cue) #C mineralization
     upt_n <- mic_n * vmax_upt_n * (don / (km_upt_n + don)) * (o2 / (km_o2 + o2)) #microbial N uptake
     death_c <- r_death * mic_c #microbial C turnover
     death_n <- r_death * mic_n #microbial N turnover
     enz_c <- p * cue * upt_c #amount of C available for enzyme production after C allocated to mineralization
     enz_n <- q * upt_n #amount of N available for enzyme production
     eprod <- ifelse(enz_c / cn_enz >= enz_n, enz_n, enz_c / cn_enz) #if enzyme production is N-limited, then number of enzyme produced = N cost for enzyme production (1 enzyme made for every 1 N used), else enzyme production C-limited
     growth_c <- (1 - p) * upt_c * cue + enz_c - cn_enz * eprod #available C for biomass growth
     growth_n <- (1 - q) * upt_n + enz_n - eprod #available N for biomass growth
     growth <- ifelse(growth_c / cn_m >= growth_n, growth_n, growth_c / cn_m) #if microbes N-limited, then amount of microbial biomass growth equals cost of N to produce microbes, else growth is C-limited
     overflow_c <- growth_c - cn_m * growth #extra C after microbial growth goes to overflow metabolism
     nmin <- growth_n - growth #N mineralization
     dmic_c <- cn_m * growth - death_c #microbial biomass C pool, growth multiplied by C:N of microbes because growth = N cost to build a microbe, but C cost is greater
     dmic_n <- growth - death_n #microbial biomass N pool
     eloss <- r_ecloss * ec #enzyme turnover
     dec <- eprod - eloss #enzyme pool
     decom_c <- vmax_c * a * ec * sol_soc / (km_c + sol_soc) #depolymerization of soc by enzymes
     decom_n <- vmax_n * (1 - a) * ec * sol_son / (km_n + sol_son) #depolymerization of son by enzymes
     dsoc <- litter_c + death_c * mic_to_soc - decom_c
     dson <- litter_n + death_n * mic_to_son - decom_n
     ddoc <- doc_input + e + decom_c + death_c * (1 - mic_to_soc) + cn_enz / (1 + cn_enz) * eloss - upt_c #doc pool, enzyme turnover deposited here.
     ddon <- don_input + e / cn_ex + decom_n + death_n * (1 - mic_to_son) + 1 / cn_enz * eloss - upt_n
#don pool, enzyme turnover deposited here. because enzymes are not split up into separate pools within the model, the amount of C and N created by the turnover of an enzyme is calculated using the C:N of enzymes
    return(list(c(dmic_c, dmic_n, dec, dsoc, dson, ddoc, ddon)))
  })
}
temp <-approxfun(soilt, method = "linear", rule = 2) #approxfun(rep(13, 4560)) #create function for input data
mois <-approxfun(soilm, method = "linear", rule = 2) #approxfun(rep(0.3, 4560)) #create function for input data
#solve function using ode solver
return(ode(state, times, derivs, pars))
}

out <- snap(pars)
```

##Plot of output
```{r plotout, echo = FALSE}
#generate plot of results
matplot(out[,1], out[,-1], type = "l", lty = 1:(ncol(out) - 1), col = 1, xlab = "time (hours)", ylab = "mol C/m3")
legend("topright", colnames(out[,-1]), lty = 1:(ncol(out) -1))
```

##Global Sensitivity Analysis
```{r globsens, eval=FALSE}
#sensitivity of variables in sensvar to the parameters in parRanges
#looped -> but could be vectorized to save time (approx 1 user sec/parameter)
for (i in 1:nrow(parRanges)){
sR <- sensRange(func = snap, parms = pars, dist = "grid", sensvar, parRange = parRanges[i,], num = 10)
  par(mar=c(4,4,3,2), mfrow = c(2,length(sensvar)))
  p1 <- plot(summary(sR), xlab = "time (hours)", ylab = "mol C/m3", legpos = "topleft", mfrow = NULL)
  p2 <- plot(summary(sR), xlab = "time (hours)", ylab = "mol C/m3", legpos = "topleft", mfrow = NULL, quant = TRUE, col = c("lightblue", "blue"))
  mtext(outer = TRUE, line = -1.5, side = 3, text = paste("Sensitivity to", rownames(parRanges[i,])), cex = 1.25)
  box("outer", lwd = 3 , col = 2)
  }
```

```{r parset, echo = FALSE, eval=FALSE}
par(mfrow = c(1,1))
```

Sensitivity range for the combination of all parameters in `parRanges`
```{r senscomb, eval=FALSE}
SensComb <- summary(sensRange(func = snap, parms = pars, dist = "latin", sensvar = sensvar, parRange = parRanges, num = 100))

plot(SensComb, main = paste("Sensitivity of", sensvar, "to all params"), xlab = "time (hours)", ylab = "mol C/m3")
box("outer", lwd = 3, col = 2)
```


##Local Sensitivity Analysis
```{r localsens, eval=FALSE}
#reminder: sensvar is predefined and may be more than one pool
SensLoc <- sensFun(func = snap, parms = pars, sensvar = sensvar, varscale = 1)

pm <- par(mfrow = c(1,length(sensvar)))
plot(SensLoc, which = sensvar, mfrow = NULL)
par(mfrow = pm)

#pairwise relationships in param sensitivity
#takes all var(s) in sensvar into account
cor(SensLoc[,-(1:2)])
pairs(SensLoc)
```

##Stability analysis
```{r stabil, eval=FALSE}
#final concentration of pools as function of params
snap.mc <- function(pars) {
  out <- snap(pars)
  return(out[nrow(out), -1])
}

#loop like in global
for (i in 1:nrow(parRanges)){
CRL <- modCRL(func = snap.mc, parms = pars, parRange = parRanges[i,])
plot(CRL)
mtext(outer = TRUE, line = -1.5, side = 3, text = paste("Final concentration response to", rownames(parRanges[i,])), cex = 1.25)
box("outer", lwd = 3 , col = 2)
}
```

#Parameter Identifiability
```{r parident, eval=FALSE}
#reminder: sensvar is predefined and may be more than one pool
#this is repeated below in mcmc but with data
coll <- collin(SensLoc)
goodColl <- coll[coll[,"collinearity"] < 20,]
goodColl[goodColl$N == (max(goodColl$N)),]
```

#Standard Parameter Fitting
```{r parfit, eval= FALSE, eval=FALSE}
#input some dummy micc data
Data1 <- matrix (nc=2,byrow=2,data= c( 2, 0.14, 4, 0.21, 6, 0.31, 8, 0.40, + 10, 0.69, 12, 0.97, 14, 1.42, 16, 2.0, 18, 3.0, 20, 4.5, 22, 6.5, 24, 9.5, 26, 13.5, 28, 20.5, 30, 29 , 35, 65, 40, 61))
colnames(Data1) <- c("time", "micc")

#input some dummy polc data (not used)
Data2 <- matrix (nc=2,byrow=2,data= c( 2, 0.10, 4, 0.21, 6, 0.30, 8, 0.40, + 10, 0.67, 12, 0.99, 14, 1.47, 16, 2.0, 18, 3.0, 20, 4.5, 22, 6.8, 24, 9.2, 26, 13.5, 28, 20.6, 30, 29 , 35, 65, 40, 61))
colnames(Data2) <- c("time", "polc")

#try putting two datasets in:
Data3 <- as.data.frame(Data1)
Data2 <- as.data.frame(Data2)
Data3$polc <- Data2$polc

#input some long-form data
Data4<- data.frame(name = c("micc", "polc", "micc", "polc", "micc", "polc", "micc", "polc"),
              time = c(1,1,2,2,3,3,4,4), val = c(0.1, 0.21, 0.3, 0.45, 0.54, 0.67, 0.8, 0.99),
              err = c(0.11, 0.22, 0.1, 0.2, 0.4,0.43, 0.4, 0.44))

#define objective function that return model-data residuals
Objective <- function(x, parset = names(x)) {
  pars[parset] <- x
  out <- snap(pars, times.def)
  ##model cost
  return(modCost(obs = Data1, model = out))
}

#identifiability check
Coll <- collin(sensFun(func = Objective, parms = pars, varscale = 1))
Coll[Coll[,"collinearity"] > 20,]

#fit model using Levenberg-Marquardt algorithm (default)
##parameters constrainted to be > 0
Fit <- modFit(p = pars2fit, f = Objective, lower = rep(0.0, length(pars2fit)))
sumfit <- summary(Fit)
#sumfit[] ##for covariance matrix & additional info

init <- snap(pars)
pars[c(names(Fit$par))] <- Fit$par
out <- snap(pars)
Cost <- modCost(obs = Data1, model = out)

plot(out, init, xlab = "time (hours)", ylab = "mol C/m3", lwd = 2, obs = Data1, obspar = list(cex = 2, pch = 18))
legend("bottomright", lwd = 2, col = 1:2, lty = 1:2, c("fitted", "original"))

plot(Cost, xlab = "time (hours)", ylab = "", main = "residuals")
```

##MCMC
```{r mcmc, eval=FALSE, eval=FALSE}
#sumfit$modVariance is the initial model error variance
#1/modVariance is drawn from gamma dist

#standard deviation of jump distribution (a multi-dimensional Gaussian)
covIni <- sumfit$cov.scaled*(2.4^2/nrow(sumfit$cov.scaled)) 

#var0 = prior mean for sigma^2, wvar0 = prior accuracy
#if wvar0 = 1, equal weight is given to prior and current value
#if wvar0 = 1, then prior is ignored
#setting var0 = sumfit#modVariance is not independent (find independent estimates...)
print(system.time(mcmc <- modMCMC(p = coef(Fit), f = Objective, jump = covIni, var0 = sumfit$modVariance, wvar0 = 1, niter = 1000, updatecov = 1001)))

plot(mcmc, Full = TRUE)

summary(mcmc)

if (length(coef(Fit)) < 2) hist(mcmc) else pairs(mcmc)

mc <- as.mcmc(mcmc$pars)
cumuplot(mc)
cov(mcmc$pars)
covIni
```

