---
title: "fmeDMC"
author: "rza"
date: "October 31, 2015"
output: html_document
---
Libraries
```{r}
library(zoo)
library(FME)
```


Set up first Obs set
```{r data1}
#Read in data from field measurements
df <-read.csv("/Users/rzabramoff/Documents/records/dropbox backup/dropbox.6.1.2015/BU/Dissertation_Research/data 2014/Ch4/mydataFlux.csv", sep=",",header=T, na.strings="NA") #scale is mgC m2 h-1 ...assume to 10cm depth?
df$SoilM <-  ifelse(df$SoilM > 0.5, 0.5, df$SoilM)

#conversions
m2cm3 <- 100*100*10

#Interpolate (linearly) gaps in data
dz <- zoo(df)
dz <- na.approx(dz)
dz <- dz[4:length(dz$scale)-1]

m <- dz$SoilM
x <- dz$SoilT #deg C
y <- dz$scale/(m2cm3) #mgC cm-3 h-1 to 10cm

##Output hourly gapfilled data object for matlab sensitivity analysis
a <- dz$SoilM
b <- dz$SoilT #deg C
c <- dz$scale
hfHourlyGapfillMatlab <- as.data.frame(cbind(a,b,c))
write.csv(hfHourlyGapfillMatlab, file = "/Users/rzabramoff/Documents/MATLAB/microbial_model_trunk/sensitivity_analysis/hfHourlyGapfilled.csv")

Obs <- as.data.frame(cbind(m,x,y)) #hourly (4556) timestep!
Obsd <- NULL
Obsd$m <-rep(tapply(dz$SoilM, dz$DOY, mean) , each = 24)
Obsd$x <-rep(tapply(dz$SoilT, dz$DOY, mean) , each = 24) 
Obsd$y <-rep(tapply(dz$scale/(m2cm3), dz$DOY, mean) , each = 24) #mgC cm-3 h-1 to 10cm
Obsd <- as.data.frame(Obsd) #hourly (4560) but no diurnal variation!

DailyGapfillForMatlab <- NULL
DailyGapfillForMatlab$mois <- tapply(dz$SoilM, dz$DOY, mean) 
DailyGapfillForMatlab$temp <- tapply(dz$SoilT, dz$DOY, mean) 
DailyGapfillForMatlab$scale <- tapply(dz$scale, dz$DOY, mean) #mgC m2 hr-1
DailyGapfillForMatlab <- as.data.frame(DailyGapfillForMatlab)
write.csv(DailyGapfillForMatlab, file = "/Users/rzabramoff/Documents/MATLAB/microbial_model_trunk/sensitivity_analysis/hfDailyGapfilled.csv")
write.table(DailyGapfillForMatlab, file = "/Users/rzabramoff/Documents/MATLAB/microbial_model_trunk/sensitivity_analysis/hfDailyGapfilled.txt", quote=F)

#Plot settings
par(mfrow = c(1,1))
```

Set up 2013 and 2014 data
```{r data2}
kath <- read.csv("/Users/rzabramoff/Dropbox (Personal)/bu/dissertation_research/damm-mcnip/Kath_resp.csv")
kath$Date <- as.Date(kath$Date)
kath$root <- kath$Control - kath$CorrectedFlux #root resp = control - trenched (het resp)

newdf <- as.data.frame(matrix(NA, nrow = 169, ncol = 10, dimnames=dimnames(kath[1:169,])))
newdf$Date <- as.Date(as.Date("2013-11-13"):as.Date("2014-04-30"))
repdf <- as.data.frame(matrix(NA, nrow = 26, ncol = 10, dimnames=dimnames(kath[1:26,])))
repdf$Date <- as.Date(as.Date("2013-10-06"):as.Date("2013-10-31"))

kz <- rbind(kath[1:171,],repdf,kath[198:209,],newdf,kath[210:length(kath$Date),])
kzoo <- zoo(kz[,c(1,2,4,6,7)])
kzoo[,3] <- na.approx(kzoo[,3])
kzoo[,4] <- na.approx(kzoo[,4])
kzoo[,5] <- na.approx(kzoo[,5])
kd <- as.data.frame(kzoo)
kd$Date <- as.Date(kd$Date)
kd$CorrectedFlux <- as.numeric(as.character(kd$CorrectedFlux)) 
kd$soilT_T <- as.numeric(as.character(kd$soilT_T))
kd$VSM_T <- as.numeric(as.character(kd$VSM_T))
plot(kd$Date, kd$CorrectedFlux) #trenched het resp in gC m-2 d-1
plot(kd$Date, kd$soilT_T)

#2013 and 2014 trenched data
Obs2 <- NULL
Obs2$m <- rep(kd$VSM_T , each = 24) #m = volumetric soil moisture in /
Obs2$x <-rep(kd$soilT_T , each = 24) #x = soil T in deg C
Obs2$y <-rep(kd$CorrectedFlux*1000/(24*m2cm3), each = 24) #y = mgC cm-3 h-1 (no diurnal var)
Obs2 <- as.data.frame(Obs2)

#2013 data only
Obs3 <- NULL
Obs3$m <- rep(kd$VSM_T[6:195] , each = 24) #m = volumetric soil moisture in /
Obs3$x <-rep(kd$soilT_T[6:195], each = 24) #x = soil T in deg C 
Obs3$y <-rep(kd$CorrectedFlux[6:195]*1000/(24*m2cm3), each = 24) #y = mgC cm-3 h-1 (no diurnal var)
Obs3 <- as.data.frame(Obs3)

#2014 data only
Obs4 <- NULL
Obs4$m <- rep(kd$VSM_T[371:560] , each = 24) #m = volumetric soil moisture in /
Obs4$x <-rep(kd$soilT_T[371:560], each = 24) #x = soil T in deg C 
Obs4$y <-rep(kd$CorrectedFlux[371:560]*1000/(24*m2cm3), each = 24) #y = mgC cm-3 h-1 (no diurnal var)
Obs4 <- as.data.frame(Obs4)
```

Consider splitting Obs2 into Obs2 &3
Change timestep of Obs to daily
Set up obs.all
```{r data3}
obs.all <- rbind(Obsd, Obs2) #no diurnal var
time <- 1:4560
  
dataUsed1 <- cbind(time,Obsd)
dataUsed2 <- cbind(time,Obs3)
dataUsed3 <- cbind(time,Obs4)
dataUsedAll <- cbind(1:length(obs.all),obs.all)

name <- rep("decom_c",4560)
datafme <- as.data.frame(cbind(name = name,time = time, val = Obsd$y))
datafme$time <- as.numeric(as.character((datafme$time)))
datafme$val <- as.numeric(as.character((datafme$val)))
```

Set up parameters for ODE
```{r}
p <- list(ea_up = 61.77)
```

Dynamic ODE model
```{r fme}
###########################
### FME results
###Setup and solve ode
Model <- function (p, times=seq(1,4561)) {
    derivs <- function(t,s,p) { #p = pars, s = state, t = time
        with(as.list(c(s,p)), {
              #ea_up <- 61.77
              vmax_0 <- 1.0815E11
              r <- 0.008314
              a <- 0.5
              ec <- 0.0325
              km_c <- 0.0025
              litter_c <- 0.0005
              r_death <- 0.00015
              mic_c <- 0.5
              mic_to_soc <- 0.5
              bd <- 0.8
              pd <- 2.52
              dgas <- 1.67
              o2airfrac <- 0.209
              dliq <- 3.17
              frac <- 0.000414
              km_o2 <- 0.121
              mois <- 0.3
              #temp <- 20
                    porosity = 1 - bd/pd
                    o2 <- dgas * o2airfrac * (porosity - mois)^(4/3)
                    sol_soc <- dliq * mois^3 * frac * soc
                    vmax_c = vmax_0 * exp(-ea_up / (r * (temp(t) + 273))) #calculate depolymerization kinetics
                    death_c <- r_death * mic_c #microbial C turnover
                    decom_c = vmax_c * a * ec * sol_soc / (km_c + sol_soc) * o2/(km_o2 + o2) #depolymerization of soc by enzymes
                    dsoc = litter_c + death_c *mic_to_soc - decom_c
                    ddecom_c = decom_c
              return(list(c(dsoc, ddecom_c)))
        })
    }
    s <- c(soc = 144.5986, decom_c = 0) 
    temp <- approxfun(dataUsed1$time, dataUsed1$x) #approxfun(rep(13,4560))
    #mois <- approxfun(dataUsed1$m)
    dieAntwoord <- ode(y = s, times=times, func=derivs, parms = p)
    return(as.data.frame(cbind(time = dieAntwoord[1:4560,1], decom_c = diff(dieAntwoord[,3])))) #fix length
}
out <- as.data.frame(Model(p))
head(out)
plot(out$time, out$decom_c*m2cm3)
```

Steady-state model for fitting
```{r fme}
###########################
### FME results
###Setup and solve ode
Model.ss <- function (p, x, m) { #p1=ea_up #p2=vmax_0
              #vmax_0 <- 1.0815E11
              r <- 0.008314
              a <- 0.5
              ec <- 0.0325
              km_c <- 0.0025
              litter_c <- 0.0005
              r_death <- 0.00015
              mic_c <- 0.5
              mic_to_soc <- 0.5
              bd <- 0.8
              pd <- 2.52
              dgas <- 1.67
              o2airfrac <- 0.209
              dliq <- 3.17
              frac <- 0.000414
              km_o2 <- 0.121
              #mois <- 0.3
              soc <- 144.5986
              decom_c <-0
              x = x
              m = m
                    porosity = 1 - bd/pd
                    o2 <- dgas * o2airfrac * (porosity - m)^(4/3)
                    sol_soc <- dliq * m^3 * frac * soc
                    vmax_c = p[2] * exp(-p[1] / (r * (x + 273))) #calculate depolymerization kinetics
                    death_c <- r_death * mic_c #microbial C turnover
                    y = vmax_c * a * ec * sol_soc / (km_c + sol_soc) * o2/(km_o2 + o2) #depolymerization of soc by enzymes
                    dsoc = litter_c + death_c *mic_to_soc - y
              return(data.frame(cbind(x, m, y))) #x=temp, m=soil moisture, y=decom_c
}

x <- dataUsed1$x
m <- dataUsed1$m
p <- c(62,1e11)    #####set par value
out <- Model.ss(p, x, m) 
head(out)
# soc <- NULL
# for (i in 1:length(out$dsoc)){
#   soc[1] <- 144.5986
#   soc[i+1] <- soc[i] + out$dsoc[i]
# }
# plot(soc)
plot(out$y)
```

FME fitting with vars
```{r}
library(FME)
# dataUsed <- as.data.frame(rbind(cbind(rep("m",length(dataUsed1[,2])), dataUsed1[,2]),cbind(rep("x",length(dataUsed1[,3])), dataUsed1[,3]),cbind(rep("y",length(dataUsed1[,4])), dataUsed1[,4])))
# names(dataUsed) <- c("name", "val")
# outfo <- Model.ss(p, dataUsed1$x, dataUsed1$m)
# outfo1 <- cbind(rep("x", length(outfo1)),outfo$x,outfo$y)
# outfo2 <- cbind(rep("m", length(outfo2)),outfo$m,outfo$y)
# outfok <- as.data.frame(rbind(outfo1,outfo2))
# names(outfok) <- c("name","x","y")
# outfok$x <- as.numeric(as.character(outfok$x))
# outfok$y <- as.numeric(as.character(outfok$y))
dataT <- dataUsed1[,c(3:4)]
dataM <- dataUsed1[,c(2,4)]

Residuals  <- function(p) {
   cost<-modCost(model=Model.ss(p, dataUsed1$x, dataUsed1$m),obs=dataT, x="x") #fix temp AND soilm
}

P      <- modFit(f=Residuals, p=c(62,1e11))
print(P$par)


plotFME <- function(){
plot(dataUsed1$x, dataUsed1$y*m2cm3,xlab="temperature",ylab="mgC m-2 h-1", pch=1, col=2, cex=1.5,
     xlim=c(0,25))
lines(Model.ss(P$par, x=0:25)$x, Model.ss(P$par, x=0:25)$y*m2cm3)
}

plotFME()
```

#Stash for now
workshop DREAM model
```{r, eval=FALSE}
Model.ssd <- function (p, x) { #p1=ea_up #p2=vmax_0
              #vmax_0 <- 1.0815E11
              r <- 0.008314
              a <- 0.5
              ec <- 0.0325
              km_c <- 0.0025
              litter_c <- 0.0005
              r_death <- 0.00015
              mic_c <- 0.5
              mic_to_soc <- 0.5
              bd <- 0.8
              pd <- 2.52
              dgas <- 1.67
              o2airfrac <- 0.209
              dliq <- 3.17
              frac <- 0.000414
              km_o2 <- 0.121
              mois <- 0.3
              soc <- 144.5986
              decom_c <-0
              x = x
                    porosity = 1 - bd/pd
                    o2 <- dgas * o2airfrac * (porosity - mois)^(4/3)
                    sol_soc <- dliq * mois^3 * frac * soc
                    vmax_c = p[2] * exp(-p[1] / (r * (x + 273))) #calculate depolymerization kinetics
                    death_c <- r_death * mic_c #microbial C turnover
                    y = vmax_c * a * ec * sol_soc / (km_c + sol_soc) * o2/(km_o2 + o2) #depolymerization of soc by enzymes
                    #dsoc = litter_c + death_c *mic_to_soc - y
              return(y) #y=decom_c
}

head(Model.ssd(p,x))
```


DREAM Calibrate 
#Sucks at higher than one dimension
```{r dreamcalibrate, eval=FALSE}
##########################
## DREAM results
unloadNamespace("dream")
library(dream)

pars <- list(p1=c(0,100),p2=c(1e10,1e15))

control <- list(
                nseq=4
                )

set.seed(456)
ptm <- proc.time() #five minutes?? what the hell... and adding temp breaks it
dd <- dreamCalibrate(FUN=Model.ssd,
                     pars=pars,
                     obs=dataUsed1$y,
                     FUN.pars=list(x=dataUsed1$x),
                     control=control
                     )
proc.time() - ptm

print(dd)

print(summary(dd))

chooseMeth = "uni.mode" #sample.ml is the default, and it does not work (at least not for fitting 1 param)
print(coef(dd, method = chooseMeth)) 

plot(dd, interactive = F)

plotFME()
lines(m2cm3*predict(dd, method=chooseMeth,newdata=list(x=0:25)),col="green")
legend("topleft",c("Data", "FME", "DREAM"), pch = c(1,NA,NA), lty = c(NA,1,1), col = c(2,1,"green"))


## Compare likelihood function for coefficients obtained by dream and FME modfit
dd$lik.fun(coef(dd, method = chooseMeth)) #dream
dd$lik.fun(P$par)    #FME

########################
## Calculate bounds around output estimates

plotCIs <- function(x,cis,...){
  em <- strwidth("M")/2
  segments(x0=x,y0=cis[,1],
           x1=x,y1=cis[,2],
           ...
           )
  segments(x0=x-em,y0=cis[,1],
           x1=x+em,y1=cis[,1],
           ...
           )
  segments(x0=x-em,y0=cis[,2],
           x1=x+em,y1=cis[,2],
           ...
           )
}##plotCIs

```

DREAM calibrate with one year, extend prediction to subsequent years
```{r dreamcalibratedata, eval=FALSE}
## Calibrate with dataUsed1
# dd <- dreamCalibrate(
#                      FUN=Model.ssd,
#                      pars=pars,
#                      obs=dataUsed1$y,
#                      FUN.pars=list(
#                        x=dataUsed1$x
#                        ),
#                      control = control
#                      )

plotFMEdat <- function(){
plot(dataUsed1$x, dataUsed1$y*m2cm3,xlab="temperature",ylab="mgC m-2 h-1", pch=1, col=1, cex=1.5,
     xlim=c(0,25), ylim=c(0,300))
points(dataUsed2$x, dataUsed2$y*m2cm3, pch = 1, col = 2)
points(dataUsed3$x, dataUsed3$y*m2cm3, pch = 1, col = 6)
lines(Model.ss(P$par, x=0:25)$x, Model.ss(P$par, x=0:25)$y*m2cm3)
}

##DataUsed1
plotFMEdat()
lines(sort(dataUsed1$x),sort(predict(dd, method = chooseMeth)*m2cm3),col="blue")
lines(sort(dataUsed1$x),sort(predict(dd,method="mean")*m2cm3),col="red")
lines(sort(dataUsed1$x),sort(predict(dd,method="median")*m2cm3),col="orange")
plotCIs(sort(dataUsed1$x),cbind(sort(predict(dd,method="CI")[,1]*m2cm3), sort(predict(dd,method="CI")[,2]*m2cm3)),col="black")

##DataUsed2
plotFMEdat()
lines(sort(dataUsed2$x),sort(predict(dd, method = chooseMeth)*m2cm3),col="blue")
lines(sort(dataUsed2$x),sort(predict(dd,method="mean")*m2cm3),col="red")
lines(sort(dataUsed2$x),sort(predict(dd,method="median")*m2cm3),col="orange")
plotCIs(sort(dataUsed2$x),cbind(sort(predict(dd,method="CI")[,1]*m2cm3), sort(predict(dd,method="CI")[,2]*m2cm3)),col="black")

##DataUsed3
plotFMEdat()
lines(sort(dataUsed3$x),sort(predict(dd, method = chooseMeth)*m2cm3),col="blue")
lines(sort(dataUsed3$x),sort(predict(dd,method="mean")*m2cm3),col="red")
lines(sort(dataUsed3$x),sort(predict(dd,method="median")*m2cm3),col="orange")
plotCIs(sort(dataUsed3$x),cbind(sort(predict(dd,method="CI")[,1]*m2cm3), sort(predict(dd,method="CI")[,2]*m2cm3)),col="black")

### Example with new sample
# dd.sim <- simulate(dd)
# plotFMEdat()
# lines(sort(dataUsed3$x),sort(predict(dd,list(x=dataUsed3$x))),col="blue")
# lines(sort(dataUsed3$x),sort(predict(dd.sim,list(x=dataUsed3$x))),col="purple")
```

DREAM MCMC
```{r dreammcmc, eval=FALSE}
dd <- dream(
                     FUN=Model.y,
                     func.type = "logposterior.density",
                     pars=pars,
                     measurement=list(y= dataUsed1$y),
                     FUN.pars=list(
                       x=dataUsed1$x
                       ),
                     control = control
                     )
```

